---
title: "Methods 4: Assignments Week 3"
output: html_notebook
---


### Problems from Chapter 4 of *Statistical Rethinking*.

#### Problem 4E1

The first line: $y_i \sim \text{Normal}(\mu,\sigma)$.

#### Problem 4E2

Two parameters, $\mu$ and $\sigma$.

#### Problem 4E3

$$
\text{Pr}(\mu,\sigma|y) = \frac{\text{Normal}(y|\mu,\sigma)\text{Normal}(\mu|0,10)\text{Exponential}(\sigma|1)}{\int\int\text{Normal}(y|\mu,\sigma)\text{Normal}(\mu|0,10)\text{Exponential}(\sigma|1)\mathrm{d} \mu \mathrm{d} \sigma}
$$

#### Problem 4E4

The second line: $\mu_i = \alpha + \beta x_i$.

#### Problem 4E5

Three: $\alpha$, $\beta$, and $\sigma$. Note that $\mu_i$ is not an independent parameter since it is a deterministic function of $\alpha$, $\beta$, and $x_i$.

#### Problem 4M1

```{r}
n <- 1e4

mus <- rnorm(n, 0, 10)
sigmas <- rexp(n, 1)
ys <- rnorm(n, mus, sigmas)
summary(ys)
```


```{r}
plot(density(ys))
```

```{r}
library(ggplot2)
ggplot(data.frame(ys), aes(ys)) +
    geom_density()
```


#### Problem 4M2

```{r}
flist <- alist(
    y ~ dnorm(mu, sigma),
    mu ~ dnorm(0, 10) ,
    sigma ~ dexp(1)
)
```


#### Problem 4M3

\begin{align}
  y_i &\sim \text{Normal}(\mu_i,\sigma) \\
  \mu_i &= \alpha + \beta x_i \\
  \alpha &\sim \text{Normal}(0, 10) \\
  \beta &\sim \text{Uniform}(0, 1) \\
  \sigma &\sim \text{Exponential}(1)
\end{align}

#### Problem 4M4

\begin{align}
  h_{ij} &\sim \text{Normal}(\mu_{ij},\sigma) \\
  \mu_{ij} &= \alpha_j + \beta x_i \\
  \alpha_j &\sim \text{Normal}(80, 10) \\
  \beta &\sim \text{Normal}(0, 10) \\
  \sigma &\sim \text{Exponential}(1)
\end{align}

$i \in \{-1, 0, 1\}$ indexes years, $j \in \{1,\ldots,n\}$ indexes students. $\alpha_j$ is the mean height (in cm) of students in the age rage considered, and $\beta$ is growth in cm/year.

#### Problem 4M5

A log-Gaussian prior for $\beta$ is more appropriate. Let's assume a prior mean growth of 8 cm/year with a standard deviation of 4 cm/year (i.e., a variance of 16 cm$^2$/year$^2$). [Looking up](https://en.wikipedia.org/wiki/Log-normal_distribution) the relevant quantities for the log-Gaussian distribution, we find:

\begin{align}
8 &= \exp \left(\mu + \frac{\sigma^2}{2}\right) \\
16 &= \left( \exp \sigma^2 - 1 \right) \exp\left(2 \mu + \sigma^2 \right)
\end{align}

After some algebra (which [a computer](https://live.sympy.org/) can do for you), we get

\begin{align}
\mu &= 4 \ln 2 - \frac{1}{2} \ln 5 \approx 1.97 \\
\sigma^2 &= \ln 5 - 2 \ln 2 \approx 0.223
\end{align}

With $\sigma \approx0.472$, we choose as our prior for $\beta$

$$
\beta \sim \text{Log-Normal}(1.97, 0.472)
$$
Let's do some simulation and see whether we like our prior.

```{r}
betas <- rlnorm(1e4, 1.97, 0.472)

ggplot(data.frame(betas), aes(betas)) +
    geom_density() +
    geom_function(fun = dlnorm, args = list(meanlog = 1.97, sdlog = 0.472), colour = "red") +
    labs(x = expression(beta),
         y = "density")
```
That looks fine.

#### Problem 4M6

The simplest way to take account of the fact the variance of heights never goes beyond 64 (meaning the *standard deviation* never goes beyond 8) is to use

$$
\sigma \sim \text{Uniform}(0, 8)
$$

### Bonus! 😃

#### Problem 4H1

We first fit the relevant model again.

```{r}
## R code 4.42
library(rethinking)
```


```{r}
# load data again, since it's a long way back
data(Howell1); d <- Howell1; d2 <- d[ d$age >= 18 , ]
```


```{r}
# define the average weight, x-bar
(xbar <- mean(d2$weight))
```


```{r}
# fit model
m4.3 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*( weight - xbar ) ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d2 )
```

Then we create a data frame containing the weights we want predictions for.

```{r}
newdata <- data.frame(weight = c(47.0, 43.7, 64.8, 32.6, 54.6))
```

We generate the posterior predictive distributions.

```{r}
heights <- sim(m4.3, data = newdata, n = 1e4)
```

The summary function gives us a quick plausibility check.

```{r}
summary(heights)
```

Looks all good. So we calculate the point predictions for the heights...

```{r}
(heights.mean <- apply(heights, 2, mean))
```

... and the 89% intervals.

```{r}
(heights.PI <- apply(heights, 2, PI, prob=0.89))
```